---
title: "Statistical Inference Project Part 2"
author: "Jing Wei Chan"
date: "24 August 2015"
output: html_document
---

## Overview

In the second portion of the project, we are going to analyze the ToothGrowth data in the R datasets package. 

## 1. Load the ToothGrowth data and perform some basic exploratory data analyses.

```{r}
library(datasets)
toothgrowth <- ToothGrowth
library(ggplot2)
ggplot(data = toothgrowth, aes(x = as.factor(dose), y = len, fill = supp)) +
    geom_bar(stat = "identity",) +
    facet_grid(. ~ supp) +
    xlab("Dose in Milligrams") +
    ylab("Tooth Length") +
    guides(fill = guide_legend(title="Supplement Type"))
```

As shown, there is a positive relationship between dose and tooth length for both OJ and VC supplement types.

## 2. Provide a basic summary of the data.

```{r}
summary(toothgrowth)
```

## 3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.

We will analyze tooth growth by carrying out t-tests at 95% confidence intervals for the 2 supplements by dosage level.

```{r}
t.test(len ~ supp, toothgrowth[toothgrowth$dose == 0.5, ])
```

At a 0.5 milligram dose, the null hypothesis is rejected as zero does not fall within the confidence interval and p-value < 0.05.

```{r}
t.test(len ~ supp, toothgrowth[toothgrowth$dose == 1, ])
```

At a 1 milligram dose, the null hypothesis is rejected as zero does not fall within the confidence interval and p-value < 0.05.

```{r}
t.test(len ~ supp, toothgrowth[toothgrowth$dose == 2, ])
```

At a 2 milligrams dose, the null hypothesis is not rejected as zero falls within the confidence interval and p-value > 0.05.

## 4. State your conclusions and the assumptions needed for your conclusions. 

From the exploratory analysis, we can see that there is a positive relationship between dose and tooth length for both OJ and VC supplement types.

From the t-tests, we can conclude that for 0.5 milligram and 1 milligram, OJ is more effective than VC in promoting tooth growth. However, for 1 milligram dose, we are unable to conclude that OJ is more effective than VC.

